Sketch Recogniton using Particle Swarm Algorithm

### 2009-04-21 draft/System.tex
after spelling correction s

\begin{equation}
\label{eq:direction}
	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {p_1 p_2 }  \cdot \overrightarrow {p_2 p_3 } }}{{\left\| {\overrightarrow {p_1 p_2 } } \right\| \times \left\| {\overrightarrow {p_2 p_3 } } \right\|}}} \right)
\end{equation}

  The direction is calculated as the angle between two vectors as in equation \ref{eq:direction} where $\overrightarrow {p_i p_j }$ is the vector from point $pi$ to point $pj$ and $\|{\overrightarrow {p_i p_j }}\|$ is the normal of the  vector $\overrightarrow{p_i p_j }$. The curvature is estimated using more than one method the result of each curvatures estimation is reported in section \ref{sec:Curvatureestimationresults}. The best curvature estimation is used, the estimation define curvature as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$.

 %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
 %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
 All these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points. Figures \ref{fig:speed2Distance} \&  \ref{fig:curvatures} show the computed information's for the stroke drawn in Figure \ref{fig:orignalStroke}.% After computing the graphs the system  compute the average, minimum and maximum values of all curves.  \\% the lower speed points correspond to vertex and dominant point's location. The higher the direction and curvature data correspond to location with higher change in curvature witch promote the location for vertices'. \\
%Description of input data. 
%how to calculate speed, curvature, area, bounding box.
%how to remove noise. 
%Finally how to compute primarily dominant points.

\begin{figure}[]
	\centering
		\includegraphics[scale=0.8]{images/orignalStroke.eps}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:orignalStroke}
\end{figure}

\begin{figure}
	\centering
			\subfigure[ Speed Graph]{	\includegraphics[scale=0.5]{images/speed2.eps}}
			\hfill
			\subfigure[ Time Difference Graph] {\includegraphics[scale=0.5]{images/time2.eps}}
	\caption{Speed and Time Difference Graphs}
	\label{fig:speed2Distance}
\end{figure}

@@

\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
\label{sec:PolygonDivisionAlgorithm}
The first algorithm, the algorithm tries to segment the stroke as a polygon. The final solution is a set of line segments that best define the input stroke. Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end. The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
 where $M$ is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}. Figure \ref{fig:DPSOERROR} shows a graphical representation of the error. The fitness is computed using equation \ref{eq:fitnessSwarm1} 
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.  If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.  This fitness function optimize two goals the first is to move the solution into the feasible solution space with acceptable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices'\cite{PolygonApproximationPSO}. 
\begin{figure}
	\centering
  \includegraphics[scale=0.8]{images/pso1.eps}			
	\caption{AlgS1 Error}% a) Possible dominant point b) Particle encoding  
	\label{fig:DPSOERROR}
\end{figure}
 %The error is chosen to in favor of larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices'.\\ %we will want to lower the number of vertices'. \\
%if we say that s
%Alg1:
%Alg2: 
\subsubsection{Hybird Curve Algorithm \textsl{AlgS2}}
\label{sec:PolygonDivisionAlgorithm}

@@

\subsection{Choosing the Best Fit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error($E_{alg}$ in Equation \ref{eq:bestFit2})represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to  compare the two segmentation fairly. 
\begin{equation}
\label{eq:bestFit1}
E_{area}=A_{st}-A_{sg}
 \end{equation}
 \begin{equation}
\label{eq:bestFit2}
E_{tot}=E_{area}+E_{alg}
 \end{equation}
 where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error. The segmentation which cross ponds to the minimum $E_{tot}$ is the best segmentation.

@@

The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.

\begin{description}
 \item [Structural and geometrical Features (FS1)] Features defines the structure of the geometrical symbol.  
\begin{description}
	 \item[Segments] Number of segments in the symbol.
	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item [Curves] Number of curves or ellipses in the symbol. 
		\item [Lines] Number of lines in the symbol. 
		\item [Perpendicular lines] Number of perpendicular lines. 
		\item [Parallel lines] Number of parallel lines. 
		\item [Intersections] Number of intersection between lines and curves. 
		\item [T intersections] Number of T intersections. 
		\item [L intersections] Number of L intersections. 
		\item [X intersections] Number of X intersections.

		%\item[No. of holes] Number of holes in the symbols.

\end{description}
\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}

\item[Statistical Features (FS3)]   
\begin{description}
	\item[Zernike moments] Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
\end{description}
	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. These features represent the global shape of the symbol drawn.
	\begin{description}
	\item [Size Ratio] Ratio between width to height of the symbol.
	\item [Ink density] Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item[Convex Hull Area] Area of convex hull with respect to area of bounding box of symbol.
	\item[Convex Hull Perimeter] Perimeter of convex hull with respect to total length of symbol.
		\item [Mean Centroidal radius] The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.

	\item [Mean Time difference] The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]
  \end{description}

\end{description}
%
%\begin{longtable}{|l|l|p{6cm}|p{5cm}|}
%	%\caption{Feature Table}
%	%\label{tab:FeatureTable}
%\hline 
%Type  & Index  & Features  & Notes \tabularnewline
%\endhead
%\hline 
%Type  & Index  & Features  & Notes \tabularnewline
%\endfirsthead
%\hline 
%Structural & \multicolumn{1}{c||}{} &  & \tabularnewline
%\hline 
% & F1.  & No. of parallel lines  & \tabularnewline
%\hline 
% & F2.  & No. of perpendicular lines  & \tabularnewline
%\hline 
% & F3.  & No. of intersection between of any two segmetns  & \tabularnewline
%\hline 
% & F4.  & No of primitives & \tabularnewline
%\hline 
% & F5.  & No. of line segments & \tabularnewline
%\hline 
% & F6  & No. of ellipse, circular arcs.  & \tabularnewline
%\hline 
%Statistical &  &  & \tabularnewline
%\hline 
% & F7.  &  & \tabularnewline
%\hline 
% & F8.  &  & \tabularnewline
%\hline 
%Rubine Features  &  & From features F9. till F13  & Introduced by in \cite{gestureexample12} \tabularnewline
%\hline 
% & F9.  & Cosine of starting angle  & \tabularnewline
%\hline 
% & F10.  & Sine of starting angle & \tabularnewline
%\hline 
% & F11.  & Length of diagonal of bounding box & (gives an idea of the size of the bounding box) \tabularnewline
%\hline 
% & F12.  & Angle of diagonal & gives an idea of the shape of the bounding box (long, tall, square) \tabularnewline
%\hline 
% & F13.  & Distance from start to end  & \tabularnewline
%\hline 
% & F14.  & Cosine of ending angle  & \tabularnewline
%\hline 
% & F15.  & Sine of ending angle  & \tabularnewline
%\hline 
% & F16.  & Total stroke length  & \tabularnewline
%\hline 
% & F17  & Change in Rotation( Arctan ) & gives the directional angle \tabularnewline
%\hline 
% & F18  & Absolute rotation  & \tabularnewline
%\hline 
% & F19  & Rotation squared  & \tabularnewline
%\hline 
% & F20  & The maximum speed reached (squared)  & \tabularnewline
%\hline 
% & F21  & Total time of stroke  & \tabularnewline
%%\newpage
%\hline 
%Zernike moments  &  &  & \tabularnewline
%\hline 
% & F22  & Zernike moments to the order  $n$  & used by \cite{zernike61} \tabularnewline
%Composite  &  &  & \tabularnewline
%\hline 
% &  &  & \tabularnewline
%\hline 
% &  &  & \tabularnewline
%\hline 
% &  &  & \tabularnewline
%\hline
%\end{longtable}
%\end{table}
  % Rubine el al.\cite{gestureexample12}  used a set of statistical features,  There are a lot of features used in the sketch recognition literature
%How to handle multi-stroked symbols. 
%The clustering algorithm.  
%The classifier used in recognition
%Describe the features used in recognition [ref to \cite{zernike61}]

 Section \ref{sec:recognitionAlgorithms} presents a comparative result of the different feature sets from these features.

\section{Classification}%%Feature Set
\label{sec:Classification}
 The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. The system uses OVO classifiers structure (object versus object). A validation algorithm is used to choose the Kernel variables parameters $c$ The results Section \ref{sec:recognitionAlgorithms} demonstrates the results of both classifiers. \\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points. . The Rubine algorithm (Linear Gaussian classifier) was implemented for comparing the system.

%%%%%%%%%%%%%%%%%%%55555555
%Symbol Recognition 16
%4.1 Preprocessing and segmentation . . . . . . . . . . . . . . . . . . . . . . . . . . . 16
%4.2 Feature-Based Statistical Symbol Recognition . . . . . . . . . . . . . . . . . . . . 18
%4.2.1 Feature Set . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 19
%4.2.2 Training . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 21
%4.2.3 Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 23
%4.2.4 User Study . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 24
%4.3 Graph-Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 28
%4.3.1 Previous Work . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 28
%4.3.2 Statistical Graph Definitions and Error-driven Stochastic Graph Matching . 29
%4.4 Image Based Symbol Recognition . . . . . . . . . . . . . . . . . . . . . . . . . . 30
%4.4.1 User Study . . . . . . . . . . . . . .
%3.3 System Components 43
%4 Segmenting a Freehand Sketch
%4.1 Interface 51
%4.2 Spline representation 52
%4.3 Points of potential feature separation 59
%4.4 Corner detection 59
%4.5 Intersection detection 61
%4.6 Clustering of interesting points 65
%4.7 Feature Extraction

>>>>

\begin{figure}
	\centering
		\includegraphics{images/curvatureCal.jpg.eps}
	\caption{Direction Calculation}
	\label{fig:curvatureCal.jpg}
\end{figure}

  The direction is calculated as the angle between vectors $Q_i$ and the $x-axis$ (vector $O_i$) as in equation \ref{eq:direction} (Figure \ref{fig:curvatureCal.jpg} explains direction calculations).We used a n estimation of curvature that define curvature as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$.

 \begin{equation}
\label{eq:direction}
	\vartheta  = \cos ^{ - 1} \left( {\frac{{\overrightarrow {Q_i}  \cdot \overrightarrow {Px_i } }}{{\left\| {\overrightarrow {Q_i} } \right\| \times \left\| {\overrightarrow {Px_i} } \right\|}}} \right)
\end{equation}
where $Q_i=\overrightarrow {P_{i - 1} P_i }$ is the vector from the point  $P_{i - 1}$ to point $P_i$ and $Px_i$ represent the $x-axis$ at this point (Figure \ref{fig:curvatureCal.jpg}).  $\|{\overrightarrow {Q_i }}\|$ is the normal of the  vector $\overrightarrow{Q_i}$

 %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
 %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
 All these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points. Figures \ref{fig:speed2Distance} \&  \ref{fig:curvatures} show the computed information's for the stroke drawn in Figure \ref{fig:AnotherorignalStroke}.% After computing the graphs the system  compute the average, minimum and maximum values of all curves.  \\% the lower speed points correspond to vertex and dominant point's location. The higher the direction and curvature data correspond to location with higher change in curvature witch promote the location for vertices'. \\
%Description of input data. 
%how to calculate speed, curvature, area, bounding box.
%how to remove noise. 
%Finally how to compute primarily dominant points.

\begin{figure}[]
	\centering
		\includegraphics[scale=0.8]{images/stroke3.jpg.eps}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:orignalStroke}
\end{figure}

\begin{figure}
	\centering
			\subfigure[ Speed Graph]{	\includegraphics[scale=0.5]{images/vel3orginal.png.eps}}
			\hfill
			\subfigure[ Time Difference Graph] {\includegraphics[scale=0.5]{images/td3.jpg.eps}}
	\caption{Speed and Time Difference Graphs}
	\label{fig:speed2Distance}
\end{figure}

\begin{figure}[]
	\centering
		\includegraphics[scale=0.8]{images/orignalStroke.eps}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:AnotherorignalStroke}
\end{figure}

@@

\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
\label{sec:PolygonDivisionAlgorithm}
The first algorithm, the algorithm tries to segment the stroke as a polygon. The final solution is a set of line segments that best define the input stroke. Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end. The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
 where $M$ is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}. Figure \ref{fig:DPSOERROR} shows a graphical representation of the error. The fitness is computed using equation \ref{eq:fitnessSwarm1} 
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.  If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.  This fitness function optimize two goals the first is to move the solution into the feasible solution space with acceptable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices'\cite{PolygonApproximationPSO}. 
\begin{figure}
	\centering
  \includegraphics[scale=0.8]{images/pso1.eps}			
	\caption{AlgS1 Error}% a) Possible dominant point b) Particle encoding  
	\label{fig:DPSOERROR}
\end{figure}
 %The error is chosen to in favor of larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices'.\\ %we will want to lower the number of vertices'. \\
%if we say that s
%Alg1:
%Alg2: 
\subsubsection{Hybrid Curve Algorithm \textsl{AlgS2}}
\label{sec:PolygonDivisionAlgorithm}

@@

\subsection{Choosing the Best Fit}
\label{sec:bestFit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to  compare the two segmentation fairly. The system choose the minimum segmentation as the final segmentation of this stroke. 
% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.
\subsection{Structural and geometrical Features(FS1)}
Features defines the structure of the geometrical symbol.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.

		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features that are composites of different geometrical and statistical symbol. These features represent the global shape of the symbol drawn.

 	\begin{itemize}
\item \emph{Size Ratio} Ratio between width to height of the symbol.
	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.

	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]
  \end{itemize}

 Section \ref{sec:featuresComparisions} presents a comparative result of the different feature sets from these features.

\section{Classification}%%Feature Set
\label{sec:Classification}
 The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. SVM is a binary classifier that maximize the margin between the decision boundary and the support vectors. To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier. This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes. The final result is based on a number of votes each shape has from all classifiers. We used the a library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.
